---
title: "Life Expectancy Final Report"
subtitle: "Data Science 2 with R (STAT 301-2)"
author: "Alice Meng, Nina Meng, Stuart Sumner, Emiliano Ghislieri"

format:
  html:
    toc: true
    toc-depth: 4
    embed-resources: true
    echo: false
    link-external-newwindow: true
    
execute:
  warning: false
  message: false
  
from: markdown+emoji 
---

## Introduction

For our final project, our group is predicting life expectancy of countries based on variables such as economic conditions and health outcomes. 

Our first recipe was an all encompassing kitchen sink model. 


```{r}
# Load package(s) ----
library(tidymodels)
library(tidyverse)


# handle common conflicts 
tidymodels_prefer()
load("datatable1.rda")
load("datatable2.rda")

```

## Kitchen Sink Results

```{r}
temp
```

Our second recipe was the same as the first recipe, with the addition of `step_corr`, with a threshold of 0.5. As we can see in the results table below, the models performed significantly worse.
```{r}
temp2
```
This solidifies our choice to go with recipe 1. Because some of the RMSEs are very similar, we will be using run times to make a further decision. 
```{r}
time_df %>% 
  DT::datatable()
```

