---
title: "Life Expectancy Progress Memo 1"
subtitle: "Data Science 2 with R (STAT 301-2)"
author: "Alice Meng, Nina Meng, Stuart Sumner, Emiliano Ghislieri"

format:
  html:
    toc: true
    toc-depth: 4
    embed-resources: true
    echo: false
    link-external-newwindow: true
    
execute:
  warning: false
  message: false
  
from: markdown+emoji 
---

``` {r}

set.seed(4444)

library(ggcorrplot)
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
load("initial_setup/tuning_setup.rda")

```

## Link to Github Repo
[link](link)

## Explanation of the Prediction Problem
We will be predicting life expectancy for the year 2015 using data from life expectancies of previous years (2000-2014) as well as predictors that can affect life expectancy, such as region, illess prevalence, and economic status. This will be a regression problem, with data from INSERT DATA LINK & details 

## Overview of the Dataset
There are a total of 3043 observations with 24(update) predictor variables. The variable types include numeric for variables such as adult mortality, Hepatitis-B immunization rates, and schooling, as well as factor variables for geographic region and binary variables for economic status (whether the countries are developed or developing). 

The target variable will be life expectancy of 2015 (the following year). 

[Describe any special issues or challenges that occurred when collecting or forming your dataset] (@Stuart do you wanna talk about this one since I think you could describe it in more detail than i could)

### Missingness Assessment

```{r}

life_train %>% 
  skimr::skim_without_charts(following_life_expect)
```

### Collinearity Assessment

``` {r}

corrplotdata = life_train %>% 
  select(-1, -19)

corr = round(cor(corrplotdata), 1)
ggcorrplot(corr)
```


### Target Variable Distribution

``` {r}

life_train %>% 
  ggplot(aes(following_life_expect)) +
  geom_histogram() +
  theme_minimal()
```