---
title: "Life Expectancy Progress Memo 2"
subtitle: "Data Science 2 with R (STAT 301-2)"
author: "Alice Meng, Nina Meng, Stuart Sumner, Emiliano Ghislieri"

format:
  html:
    toc: true
    toc-depth: 4
    embed-resources: true
    echo: false
    link-external-newwindow: true
    
execute:
  warning: false
  message: false
  
from: markdown+emoji 
---

```{r}
# set seed
set.seed(4444)
# load packages + data
library(ggcorrplot)
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
load("initial_setup/tuning_setup.rda")
load("results/null_fit.rda")

```

## Link to Github Repo

[Github Repo Link](https://github.com/STAT301-3-2023SP/final-project-datatists)

## Basic Recipes
The first recipe is a kitchen sink which encompasses all of the variables. The second 
recipe includes `step_corr` to remove variables that have high correlation with each other.

## Assessment Measure
RMSE will be used for model comparison and selection. However, our overall metric set will include 
Rsq and MAE, and tuning time will be used if metrics are similar between two models.

## Baseline Models
Our null model had a RMSE of 9.47 years. 

```{r}
null_results <- null_tune %>% 
  collect_metrics()
null_results
```


## Issues
There were no major issues that we had when fitting our models. 